Ancestral traits estimations

### 2021-03-22 Analysis/02-ancestral_states_estimations.Rmd
added visualisation for ancestral shape spaces

```{r}
library(ape)
library(dispRity)
source("../FUnctions/ancestral.states.fun.R")
```

@@

@@

## Building the matrices
psita_max_ace <- lapply(psita_ace, extract.ace, what = "max",
                         matrix = spaces_list$psittaciformes)
psita_10n_ace <- lapply(psita_ace, extract.ace, what = 10, sampling = "normal",
                         matrix = spaces_list$psittaciformes)
chara_max_ace <- lapply(chara_ace, extract.ace, what = "max",
                         matrix = spaces_list$charadriiformes)
chara_10n_ace <- lapply(psita_ace, extract.ace, what = 10, sampling = "normal",
                         matrix = spaces_list$psittaciformes)
```

@@

>>>>

```{r}
library(ape)
library(dispRity)
source("../Functions/ancestral.states.fun.R")
```

@@

> TODO: improve the ACE speed exploring castor::asr_independent_contrasts, castor::asr_squared_change_parsimony or castor::asr_subtree_averaging

@@

## Building the matrices
psita_max_ace <- lapply(psita_ace, extract.ace, what = "max",
                         matrix = spaces_list$psittaciformes)
psita_10n_ace <- lapply(psita_ace, extract.ace, what = 10, sampling = "normal",
                         matrix = spaces_list$psittaciformes)
chara_max_ace <- lapply(chara_ace, extract.ace, what = "max",
                         matrix = spaces_list$charadriiformes)
chara_10n_ace <- lapply(chara_ace, extract.ace, what = 10, sampling = "normal",
                         matrix = spaces_list$charadriiformes)
```

@@

## Visualisation

```{r, fig.height = 18, fig.width = 12}
par(mfrow = c(3, 2))

## Loading the full space
plot.dispRity(spaces_list$psittaciformes, col = "blue", main = "Psittaciformes section of the shape space", xlab = "PC1", ylab = "PC2")
plot.dispRity(spaces_list$charadriiformes, col = "orange", main = "Charadriiformes section of the shape space", xlab = "PC1", ylab = "PC2")

## Full space with max ancestor
plot(custom.subsets(psita_max_ace[[1]], group = list("Nodes" = 342:681, "Tips" = 1:341)), col = c("grey", "blue"), pch = c(19, 19), main = "Psittaciformes with nodes (max. lik.; tree 1)", xlab = "PC1", ylab = "PC2")
plot(custom.subsets(chara_max_ace[[1]], group = list("Nodes" = 360:717, "Tips" = 1:359)), col = c("grey", "orange"), pch = c(19, 19), main = "Charadriiformes with nodes (max. lik.; tree 1)", xlab = "PC1", ylab = "PC2")

## Full space with random ancestor
plot(custom.subsets(psita_10n_ace[[1]], group = list("Nodes" = 342:681, "Tips" = 1:341)), col = c("grey", "blue"), pch = c(19, 19), main = "Psittaciformes with nodes (random; tree 1)", xlab = "PC1", ylab = "PC2")
plot(custom.subsets(chara_10n_ace[[1]], group = list("Nodes" = 360:717, "Tips" = 1:359)), col = c("grey", "orange"), pch = c(19, 19), main = "Charadriiformes with nodes (random; tree 1)", xlab = "PC1", ylab = "PC2")
```
